Validating risk scores - Czech data

### 2021-01-11 other_outputs/risk_scores_consortium.Rmd
Updates for risk scores based on feedback by V.M.T.deJong

```{r}
data_risk <- data$patient_data %>%
  mutate(mortality  = case_when(
    last_record >= 30 ~ "alive",
    outcome == "Discharged" ~ "alive",
    outcome == "Death" ~ "dead",
    TRUE ~ "alive")) %>%
  inner_join(marker_data_admission, by = c("patient_id", "hospital_id")) %>%
  mutate(creatinin = if_else(is.na(creatinin.x), creatinin.y, creatinin.x),
         pt_inr = if_else(is.na(pt_inr.x), pt_inr.y, pt_inr.x),
         albumin = if_else(is.na(albumin.x), albumin.y, albumin.x),
         ) %>%
  select(-creatinin.x, -creatinin.y, -pt_inr.x, -pt_inr.y, -albumin.x, -albumin.y) %>%
  #avoid including some derived quantities
  select(-age_norm, -high_creatinin, -high_pt_inr, -low_albumin, -heart_problems, -lung_problems, -markers_out_of_range, -antibiotics, -macrolides) %>%
  #avoid including indicators that are not available at hospitalization 
  select(-discontinued_medication, -last_record, -outcome,  -best_supportive_care_from,
         -starts_with("ever_"), -starts_with("any_"), -starts_with("worst"), -had_invasive,
         -first_day_invasive, -last_day_invasive) %>%
  #avoid columns that are not useful
  select(-hospital_id, -patient_id, -pcr_positive, -FiO2, -urea,
         -took_convalescent_plasma, -obesity, -took_favipiravir, -took_az,
         -took_dexamethasone, -took_remdesivir, -IgG, -Horowitz_index,
         -has_hypertension_drugs
         ) %>%
  #convert to factors to use with mice
  mutate(sex = factor(sex), 
         breathing_s = factor(breathing_s, ordered = FALSE), 
         requires_oxygen = breathing_s != "AA",
         across(all_of(names(.)[sapply(., typeof) == "logical"]), factor))
```

@@

block_vars <- setdiff(names(data_risk), "mortality")
blocks <- make.blocks(block_vars)
predictor_matrix <- make.predictorMatrix(data_risk, blocks)

predictor_matrix[, "requires_oxygen"] <- 0
low_obs_count <- c("lactate_dehydrogenase", "AST", "ferritin", "IL_6", "leukocyte_count", "PEEP")
predictor_matrix[low_obs_count, ] <- 0
predictor_matrix[low_obs_count, c("age","sex")] <- 1

@@

```{r}
validation_res <- list()
for(model in c("hu", "xie", "wang_clinical", "age_baseline")) {
  cat("\n===== Validating ", model, " ==============\n")
  single_res <- list()
  cat("- AUC\n")
  single_res$est_auc <- validate.mids(data = data_risk_imp,
                       model = model,
                       measure = auc,
                       quiet = F)

@@

  cat("- Calibration intercept\n")
  single_res$est_int <- validate.mids(data = data_risk_imp,
                       model = model,
                       measure = calibration_intercept)

>>>>

```{r}
data_risk <- data$patient_data %>%
  mutate(mortality  = case_when(
    last_record >= 30 ~ "alive",
    outcome == "Discharged" ~ "alive",
    outcome == "Death" ~ "dead",
    TRUE ~ "alive")) %>%
  inner_join(marker_data_admission, by = c("patient_id", "hospital_id")) %>%
  mutate(creatinin = if_else(is.na(creatinin.x), creatinin.y, creatinin.x),
         pt_inr = if_else(is.na(pt_inr.x), pt_inr.y, pt_inr.x),
         albumin = if_else(is.na(albumin.x), albumin.y, albumin.x),
         ) %>%
  select(-creatinin.x, -creatinin.y, -pt_inr.x, -pt_inr.y, -albumin.x, -albumin.y) %>%
  #avoid including some derived quantities
  select(-age_norm, -high_creatinin, -high_pt_inr, -low_albumin, -heart_problems, -lung_problems, -markers_out_of_range, -antibiotics, -macrolides, -comorbidities_sum_na) %>%
  #avoid columns that are not useful
  select(-hospital_id, -patient_id, -pcr_positive, -FiO2, -urea,
          -discontinued_medication, -last_record, -outcome,  
         #-best_supportive_care_from,
         -starts_with("ever_"), ever_hcq,
         -starts_with("any_"), 
         -starts_with("worst"), worst_breathing_s,
         -had_invasive,
         -first_day_invasive, -last_day_invasive,
      #   -startsWith("took"),
         -took_hcq,
         -took_convalescent_plasma, -obesity, -took_favipiravir, -took_az,
         -took_dexamethasone, -took_remdesivir, -IgG, -Horowitz_index,
         -has_hypertension_drugs
         ) %>%
  #convert to factors to use with mice
  mutate(sex = factor(sex), 
         breathing_s = factor(breathing_s, ordered = FALSE), 
         worst_breathing_s = factor(worst_breathing_s, ordered = FALSE), 
         requires_oxygen = breathing_s != "AA",
         across(all_of(names(.)[sapply(., typeof) == "logical"]), factor))
```

These are the final columns used:
```{r}
names(data_risk)
```

@@

# I don't use mortality as it is already included in worst_breathing_s
block_vars <- setdiff(names(data_risk), "mortality")
blocks <- make.blocks(block_vars)
predictor_matrix <- make.predictorMatrix(data_risk, blocks)

predictor_matrix[, "requires_oxygen"] <- 0
predictor_matrix[, "mortality"] <- 0
predictor_matrix["oxygen_flow", "worst_breathing_s"] <- 0
low_obs_count <- c("lactate_dehydrogenase", "AST", "ferritin", "IL_6", "leukocyte_count", "PEEP")
predictor_matrix[low_obs_count, ] <- 0
predictor_matrix[low_obs_count, c("age","sex")] <- 1

@@

```{r}
validation_res <- list()
for(model in c("hu", "xie", "wang_clinical", "age_baseline")) {
  cat("\n===== Validating ", model, " ==============\n")
  single_res <- list()
  cat("- AUC\n")
  single_res$est_auc <- validate.mids(data = data_risk_imp,
                       model = model,
                       measure = covid19retrospective::auc,
                       quiet = F)

@@

  cat("- Calibration intercept\n")
  single_res$est_int <- validate.mids(data = data_risk_imp,
                       model = model,
                       measure = calibration_intercept,
                       use_logistf = (model == "hu"))
